% ------------------------------ Related Studies -----------------------------

\section{Related Studies}

\begin{comment}
    For avoiding plagiarism, citations should be used for all referred texts particularly here and other parts of the document using appropriate numbers within square bracket for all mapped references under \textbf{References} section. You should check any standard journal paper for typical use of citations.     
\end{comment}

\noindent
The intersection of social media analytics and mental health research has received increasing attention in recent years, leading to several important studies that highlight the potential for early detection and intervention. This section reviews key findings from various studies, emphasizing the relevance and applicability of social media data for identifying mental health disorders. \\

\noindent
One of the seminal works in this domain is by Choudhury et al. (2013), who explored the predictive capabilities of social media content in identifying depression. They analyzed Twitter data and discovered that specific linguistic patterns, such as the use of negative emotion words, correlated strongly with self-reported depressive symptoms. This study demonstrated that social media could serve as a valuable resource for predicting mental health conditions, offering a potential tool for clinicians and researchers alike \cite{Choudhury2013PredictingDV}. \\

\noindent
Similarly, Guntuku et al. (2017) conducted an integrative review that focused on detecting mental illness through social media. Their work synthesized various approaches and methodologies used in the field, providing insights into the effectiveness of different machine learning algorithms and sentiment analysis techniques. They found that social media platforms are rich sources of data that can reveal critical information about users' mental health, advocating for the development of robust systems to analyze this data effectively \cite{Guntuku2017DetectingDA}. \\

\noindent
A systematic review by Mathur et al. (2022) further emphasized the significance of mental health classification on social media. They examined various studies that utilized machine learning techniques for mental health detection, highlighting the success of these models in identifying depression and anxiety based on user-generated content. Their findings reinforced the notion that social media can be leveraged not only for individual assessments but also for broader epidemiological studies to understand population mental health trends \cite{Mathur2022MentalHC}. \\

\noindent
In addition, Nadeem (2016) contributed to the discussion by investigating depression identification on Twitter. The study focused on developing algorithms that could discern emotional cues in tweets, indicating a user's mental state. The findings revealed that simple text analysis could lead to significant improvements in identifying at-risk individuals, further validating the potential of social media data in mental health monitoring \cite{nadeem2016identifying}. \\

\noindent
Research by AlSagri and Ykhlef (2020) introduced a machine learning-based approach specifically for depression detection on Twitter. Their study incorporated both content and activity features, demonstrating that a combination of linguistic and behavioral analysis could enhance the accuracy of depression identification. This work illustrated the multifaceted nature of social media data and its ability to capture not just what users say but also how they interact online \cite{alsagri2020machine}. \\

\noindent
In a more recent study, Vaishnavi et al. (2022) investigated the application of various machine learning algorithms for predicting mental health illnesses. They found that certain algorithms outperformed others in classifying mental health conditions based on social media posts. This study provided a comparative analysis that could inform future research directions, emphasizing the importance of algorithm selection in the context of mental health detection \cite{Vaishnavi_2022}. \\

\noindent
Lastly, Safa et al. (2023) presented a roadmap for future development in predicting mental health using social media. Their work highlighted the ongoing challenges in the field, including ethical considerations and the need for improved data privacy measures. They emphasized that while social media offers rich data for mental health analysis, researchers must approach this opportunity with a strong ethical framework to ensure user safety and data security \cite{safa2023predictingmentalhealthusing}. \\

\noindent
These studies collectively underscore the growing body of evidence supporting the integration of social media analytics and machine learning for mental health detection. They provide a solid foundation for the current project, which aims to enhance existing methodologies and develop a predictive model for identifying mental health disorders from social media posts.


% ----------------------------- Related Studies End --------------------------
